---
title: "Simple Linear Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
# Load your libraries
library(mosaic)
library(DT)
library(car)
library(tidyverse)
library(pander)

# Load your data after saving a csv file in your Data folder.
# You can use either 
#   someName <- read.csv("../Data/YourDataFileName.csv", header=TRUE)
# or
#   library(readr)
#   someName <- read_csv("../Data/YourDataFileName.csv")

# Don't forget to run "Session -> Set Working Directory -> To Source file location"

```




<!-- Directions`

Come up with a simple linear regression of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.

* You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.

-->


## Background

I am interested in movies and their ability to generate revenue. Movie making, though an art, often involves massive business ventures and can be an easy way to lose millions of dollars in investments. Needless to say, being able to predict movie success before the investment is made is a very important, and useful, skill. 

One of the main, assumptions up and coming film makers make is that in order for them to make good movies they need money. I'll be testing that notion in this analysis.

## Question

We'll model that question as follows:

$$
  \underbrace{Y_i}_\text{Revenue} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Budget} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2) 
$$
Our null hypothesis is that the slope of the regression model,$\beta_1$ is equal to 0, or in other words that the Budget and Revenue are not related. OUr alternative hypothesis is that the slope is not equal to 0 and, therefore, that there is some sort of relationship between the two factors.
$$
  H_0: \beta_1 = 0
$$
$$
  H_a: \beta_1 \neq 0
$$
$$
  \alpha = 0.05
$$

## The Data
For this analysis we'll use data collect by the The Movie Database. Their data has been collected for download at on Kaggle.com as well, [click here for the link.](https://www.kaggle.com/tmdb/tmdb-movie-metadata) 

## {.tabset .tabset-pills .tabset-fade}

### Hide Data

### Show Data {.tabset}

```{r, message=FALSE, warning=FALSE}
library(readr)
movies <- read_csv("../../Data/tmdb_5000_movies.csv")
```

  
The Kaggle website notes that budget values that equal zero should probably be treated as if they were null. I'll filter them out of our dataframe and do the same for revenue values.
  
```{r, warning=FALSE}
movies2 <- movies %>% 
  select(
    c(original_title, budget, revenue, release_date)
  ) %>% 
  filter(
    budget != 0 & revenue != 0
  ) %>% 
  droplevels()

```

```{r, warning=FALSE}
datatable(movies2)
```

## Analysis

For this analysis, we'll conduct test for linear regression, plot a scatterplot and a regression line, check assumptions and find the confidence intervals.
```{r Perform the Regression}
movies2.lm <- lm(
  revenue ~ budget,
  data = movies2
)
```

```{r Plot}
plot(revenue ~ budget,
     data = movies2,
     pch = 21,
     bg = "skyblue",
     col = "gray",
     main = "Movie Revenue and Budget"
     )

abline(
  movies2.lm, 
  lwd = 3,
  col = rgb(.4, .4, .4, .2)
)

abline(v=seq(40,120,20), h=seq(40,120,20), lty=2, col=rgb(.6,.6,.6,.2))

```
```{r, message=FALSE}
movies2 %>% 
  summarise(cor(budget, revenue, use = "complete.obs")) %>% 
              pander( )
```

Though the dots aren't tightly clustered, there does seem to be a consistency in the slope. The correlation coefficient is 0.705, so there does seem to be a correlation.

The estimated slope is 2.96, which is significantly different from zero since $p=0 $ for this analysis. Below I have shown the complete summary of the regression.

```{r Show the Regression Summary}
pander(summary(movies2.lm))
```

The mathematical equation for the scatterplot above is

$$
  \underbrace{\hat{Y}_i}_\text{Mean Revenue} = 902541 + 2.96 \underbrace{X_i}_\text{Budget}
$$
Since our p-value is less than our alpha we can assume that we have sufficient evidence to reject the null hypothesis. First, however let's check the linear regression assumptions.

## Appropriateness of the Regression

The Regression fails two assumptions. The resideuals in the Q-Q plot are not normal, though the Residuals vs Fitted plot shows that the data is essentially linear. The Residuals vs. Order plot shows that there the error terms seem to be correlated due to the obvious megaphone shape we see there. 
```{r Check Assumptions}
par(mfrow = c(1,3))
plot(movies2.lm, which = 1:2)
plot(movies2.lm$residuals, main = "Residuals vs Order")
```

When checking the confidence intervals below, we see that the intercept varies greatly. Revenue could be anywhere between 5.3 million in the negative to 7.1 milliion in the positive.
```{r Finding Confidence Intervals for Model Parameters}

confint(movies2.lm, level = .95) %>% pander()
```


## Conclusion
More research should be done into the way the data was gathered and perhaps this test should be conducted with inflation accounted for since the data points span from 1916 to 2017. As it stands, since we haven't met the assumptions for the analysis, we can't conclude that the more money you put into a movie production the more revenue you'll earn. However, further examination might yield more useful results.