---
title: "Predicting the Weather"
author: "Brian Grant"
date: "9/25/2020"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---
```{r Prep environment}
library(mosaic)
library(tidyverse)
library(DT)
```
```{r Clean the Data}
data <- read.csv("C:/Users/BAG20/Downloads/Temporary/Book1.csv")
data <- data %>% 
  mutate(
    day = as.numeric(substr(as.character(data$ï..Date), start = 1, stop = 2))
  ) %>%
  filter(
    Hi_temp != 'NA'
  ) %>% 
  select(c(Hi_temp, day))
mutate <- as.numeric(data$day)
```

```{r Perform the Regression}
mylm <- lm(Hi_temp ~ day, data = data)
```

```{r Graphic}
# show data points

# show predicted temperature for 21-Sep-2020
prediction <- predict(mylm, data.frame(day = 21))

# label 

```



# Background
For this analysis, we'll simply be practicing linear regression with weather data. The task is to take 8-30 data points and predict the highest temperature Rexburg, Idaho will experience on September 21, 2020.

# Conclusion
In the end, we predict the temperature wil be 70°F on September 21, 2020. We also conclude that there isn't a meaningful relationship between the date and the high temperature, however simple linear regression isn't appropriate for our data as it's assumptions can not be met in this analysis.

# Graphical Summary

```{r}
# show regression line
ggplot(data, aes(x = day, y = Hi_temp)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```
This scatterplot shows a curve instead of a linear relationship, which further confirms why simple linear regression is not appropriate for a data.

# Technical Details

Now that we've summarised the results, I'll now show you the process I followed.

## Questions & Hypotheses

Our hypotheses are as follows:

$H_0:\beta_1 = 0$
$H_a:\beta_1 \neq 0$

<!-- $H_0:\beta_0 = 0$ -->
<!-- $H_a:\beta_0 \neq 0$ -->

In other words, I am looking for sufficient evidence to conclude that there is a meaningful relationship between the average date and temperature.

# Analysis

```{r}
summary(mylm)
```


# Technical Details

- How the data was collected
It was manually added to a csv from [the following link]( https://www.timeanddate.com/weather/@5605242/historic?month=1&year=2019)

## {.tabset .tabset-pills .tabset-fade}

### Hide Data

### Show Data {.tabset}


- Show the data set

```{r}
datatable(data)
```


- show labeled regression model and hypotheses

- present lm summary, diagnostic plots, and their interpretations


```{r Check Assumptions}
par(mfrow = c(1, 3))
plot(mylm, which = 1:2)
plot(mylm$residuals)
```

- present fitted (y-hat) regression equation

- present an interpreation of estimated slope

- show the math behind the predicted temp
$$
  \underbrace{Y_i}_\text{Hi_temp} = \beta_0 + \beta_1 \underbrace{X_i}_\text{day} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2) 
$$

