---
title: "Week 10 | Simple Logistical Regression"
author: "Brian Grant"
date: "6/22/2020"
output: html_document
---
# Announcements

# Day 1
## Logistic Regression
[Simple Logistic Regression](LogisticRegression.html)

- notice that if your binomials have a value greater than 0, you can have the logistical regression have the y value set to [variable] > 0 so that it'll accept anything above 0 as a success (i.e. '1').

This week we're collecting our own data for the Logistic Regression analysis.

Notice that for this we want the p-value to be GREATER than alpha. This is because the null hypothesis is backwards than normal: we're TRYING to accept that null is true instead of trying to reject the null.

# Day 2

## Class Activity - Simple Logistic Regression (Part 2)

```{r}
library(mosaic)
```

```{r}
KidsFeet.glm <- glm(sex == "B" ~ length,
                    data = KidsFeet,
                    family = binomial)
summary(KidsFeet.glm)

coef1 <- KidsFeet.glm$coefficients
coef1

plot(sex == "B" ~ length, data = KidsFeet, col = c("blue", "green"))
curve(exp(coef1[1] + coef1[2]*x)/(1 + exp(coef1[1]+coef1[2]*x)), add = TRUE)

predict(KidsFeet.glm, newdata = data.frame(length = 0), type = "response")

exp(coef1[2])
```

# In class excercise
```{r}
#View(infert)
#?infert
infert.glm <- glm( (spontaneous > 0) ~ age, data=infert, family=binomial)

summary(infert.glm)
plot( (spontaneous > 0) ~ age, data=infert)
curve( exp(1.48706 + -0.05616*x)/(1 + exp(1.48706 + -0.05616*x)), add=TRUE)

table(infert$age)
library(tidyverse)
library(mosaic)
library(pander)    
library(plyr)

#Residual deviance: 334.01  on 246  degrees of freedom
pchisq(334.01, 
       246,
       lower.tail=FALSE)

library(mosaic)
#View(Galton)
#?Galton

Galton.glm <- glm(sex == "M" ~ height,
                  data = Galton,
                  family = binomial)
summary(Galton.glm)
coef.gal <- Galton.glm$coefficients

plot(sex == "M" ~ height,
     data = Galton)
curve(exp(coef.gal[1] + coef.gal[2]*x)/
        (1+exp(coef.gal[1]+coef.gal[2]*x)),
                                        add = TRUE)
```

# Assessment Quiz


## Q1
Run the following codes in R.



```{r}
  # library(mosaic)
  # 
  # View(Gestation)
  # 
  # ?Gestation
```

Use an appropriate test to determine if the birth weight in ounces of a baby (wt) can be used to predict the probability that the mother smoked at all (smoke>0) during or prior to the pregnancy. The graphic of the correct analysis is shown below.

```{r}
Gest.glm <- glm(smoke > 0 ~ wt,
                data = Gestation,
                family = binomial)

summary(Gest.glm)
```

What are the values of b0 and b1 for the curve shown in the graphic below?

Answer: b0=2.417 and b1=−0.018 


### Explanation

Call:
glm(formula = smoke > 0 ~ wt, family = binomial, data = Gestation)

Deviance Residuals:
    Min       1Q   Median       3Q      Max  
-1.8055  -1.2341   0.8984   1.0765   1.4810  

Coefficients:
             Estimate Std. Error z value Pr(>|z|)    
(Intercept)  2.416983   0.402280   6.008 1.88e-09 ***
wt              -0.018277   0.003311  -5.520 3.39e-08 ***

## Q2
Here is the output of a logistic regression performed in R.

Coefficients:
            Estimate Std. Error z value Pr(>|z|)   
(Intercept)  -6.6035     2.3514  -2.808  0.00498
mpg             0.3070     0.1148   2.673  0.00751

What is the predicted value for P(Yi=1|xi=25) ?

```{r}
exp(-6.6035 + 0.3070*25) / (1 + exp(-6.6035 + 0.3070*25))
```
Answer: 0.7448821

### Explanation
To compute a predicted value we can either do the calculation by hand:

exp(-6.6035 + 0.3070*25) / (1 + exp(-6.6035 + 0.3070*25)) = 0.7448821

Or we can use the predict function:

predict(glmObject, data.frame(mpg=25), type="response")

However, in this particular case, we don't have access to the "glmObject" that created the summary output, so we can only compute the answer by hand.

 

## Q3

Consider again the logistic regression output shown below.

Coefficients:
            Estimate Std. Error z value Pr(>|z|)   
(Intercept)  -6.6035     2.3514  -2.808  0.00498
mpg             0.3070     0.1148   2.673  0.00751

Which of the following correctly interprets these results


```{r}
exp(0.3070)
```

Answer: The odds that Yi=1 increase by 36% for every 1 unit change in x.

### Explanation

To intepret the effect of mpg (the x-variable) on the odds, we compute exp(0.3070), which gives the value of eb1 . See the Overview page of Logistic Regression for an explanation.

Coefficients:
            Estimate Std. Error z value Pr(>|z|)   
(Intercept)  -6.6035     2.3514  -2.808  0.00498
mpg             0.3070     0.1148   2.673  0.00751

 

Since exp(0.3070) = 1.359341, we find that the odds of a success, i.e., that Yi =1 increases by 35.9341%, which rounds to 36% for every 1 unit increase in x.

Notices that the 35.9341 came from 1.359341. If you have 1.36 of what you originally had, then you increased by 0.36 or 36%.

100% - 136% = 36%



