---
title: "W10_class_notes"
author: "Brian Grant"
date: "6/22/2020"
output: html_document
---
# Announcements

# Day 1
## Logistic Regression
[Simple Logistic Regression](LogisticRegression.html)

- notice that if your binomials have a value greater than 0, you can have the logistical regression have the y value set to [variable] > 0 so that it'll accept anything above 0 as a success (i.e. '1').

This week we're collecting our own data for the Logistic Regression analysis.

Notice that for this we want the p-value to be GREATER than alpha. This is because the null hypothesis is backwards than normal: we're TRYING to accept that null is true instead of trying to reject the null.

# Day 2

## Class Activity - Simple Logistic Regression (Part 2)

```{r}
library(mosaic)
```

```{r}
KidsFeet.glm <- glm(sex == "B" ~ length,
                    data = KidsFeet,
                    family = binomial)
summary(KidsFeet.glm)

coef1 <- KidsFeet.glm$coefficients
coef1

plot(sex == "B" ~ length, data = KidsFeet, col = c("blue", "green"))
curve(exp(coef1[1] + coef1[2]*x)/(1 + exp(coef1[1]+coef1[2]*x)), add = TRUE)

predict(KidsFeet.glm, newdata = data.frame(length = 0), type = "response")

exp(coef1[2])
```


