---
title: "Week 12 | Chi-Squared"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Prep, include=FALSE}
library(tidyverse)
library(mosaic)
library(pander)
library(car)
```

# Announcements

# Day 1

## Chi Squared Analysis Example


```{r}
#View(Titanic)
#?Titanic
```

Lets see if there was a relationship between passenger survival and ticket class.
```{r}
table(TitanicSurvival$passengerClass, TitanicSurvival$survived)
```


Lets see if there was a relationship between passenger survival and gender.

```{r}
table(TitanicSurvival$sex, TitanicSurvival$survived)
```

### how c() can be used in Chi-squared Analysis

Bascially, you bind c() lists together into a matrix.

Look up rbind & cbind. They turn c() lists into a matrix. rbind turns each list and turns it into a row. cbind does the same but turns them into columns.

```{r}
fruit <- rbind(Apples=c(Good = 80, Bruised = 20, Rotten = 15),
               Oranges=c(Good = 75, Bruised = 25, Rotten = 10)
               )
fruit

fruit <- cbind(Apples=c(Good = 80, Bruised = 20, Rotten = 15),
               Oranges=c(Good = 75, Bruised = 25, Rotten = 10)
               )
fruit
```


If you need to flip the matrix, just use the t() function.
```{r}
t(fruit)
```


### how to graphically summarise the data

Use a barplot.

```{r}
barplot(fruit, beside = TRUE, legend.text = TRUE)
```

Is type of fruit connected to it's condition?
```{r}
chisq.test(fruit)
```
We have sufficient evidence to reject the null and accept that fruit type and condition are related.
