---
title: "W3_Notes"
output: html_document
---
## Setup

```{r, include=FALSE}
library(tidyverse)
library(plotly)
library(mosaic)
library(car)
```
We were working on content in MakingInference.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pander)
library(mosaic)
library(car)
```

```{r}
View(mtcars)
```

## One sample t test
```{r}
t.test(mtcars$mpg, mu=20, alternative='two.sided',conf.level=0.95)

ttestoutput <- t.test(mtcars$mpg, mu=20, alternative='two.sided',conf.level=0.95)

pander(ttestoutput)

args(t.test)

qqPlot(mtcars$mpg)
```
```{r}
t.test(KidsFeet$length, 
       mu = 28, alternative = "two.sided", conf.level = .95)
```

## Two sample independent t test
```{r}
t.test(mpg ~ am, 
       data=mtcars, 
       mu=0)
```

```{r}
qqPlot(length ~ sex, data = KidsFeet)

t.test(length ~ sex, 
       data = KidsFeet,
       mu = 0, 
       alternative = "two.sided",
       conf.level = .95)
```


## Two sample dependent (paired samples) t test

Use the R Instructions to perform a Paired-Samples t Test of the following hypotheses using the KidsFeet data set.


```{r}
kf <- KidsFeet
kf$width3 <- kf$width * 3

```



# Skills Quiz

In which distribution would an observation of x=5 be the most likely to occur?

```{r}
curve(dnorm(x,-5,2), -12, 20, col="skyblue", lwd=2, ylab="", n=1000, ylim=c(0,.81))

curve(dchisq(x,5), from=0, add=TRUE, col="firebrick", lwd=2, n=1000)

curve(dt(x,5), add=TRUE, col="skyblue4", lwd=2, n=1000)

curve(df(x,10,10), from=0, add=TRUE, col="salmon", lwd=2, n=1000)

legend("topright",
       legend=c("Normal(-5,2)",
                "Chi-squared(8)",
                "t(5)",
                "f(2,5)"
                ),
       lwd=2,
       lty=1,
       col=c("skyblue","firebrick","skyblue4","salmon"),
       bty='n')

```
Which of the following set of sample values are most likely to be obtained if a sample of n = 5 observations is sampled from the normal distribution drawn by the above code in R?

```{r}
curve(dnorm(x, -5, 2), from=-12, to=2, lwd=2, col="skyblue", ylab="Normal Distribution with Mean of -5 and Standard Deviation of 2")

abline(h=0, v=-5, lty=2)
```

What are the mean uptake values for each Type of plant (Quebec and Mississippi) for plants that are chilled at a concentration of 250?

Use a subset of the CO2 dataset in R to answer the following questions.

Filter the dataset so that you are considering only the chilled plants where conc = 250.

(Hint: Treatment == "chilled" & conc == 250)

```{r}
?CO2

View(CO2)

library(tidyverse) #loads the filter(...) function

CO2.chilled.250 <- filter(CO2, Treatment == "chilled"& conc == 250)

CO2.chilled.250 %>% 
  group_by(Type) %>% 
  summarise(mean(uptake))
```

Recreate one of these dot plots in R using either Base or ggplot2 graphics.

```{r}
stripchart(uptake ~ Type,
           data = CO2.chilled.250,
           col = c('skyblue', 'red'),
           pch = 15,
           ylim = c(-.2, 3),
           method = "stack",
           main = "Quebec shows higher uptake values than \nMississippi when plants are chilled at an ambient CO2 of 250")
           
```

These results are based on a small sample of Mississippi and Quebec plants. The question that remains is whether or not this pattern holds for ALL Mississippi and Quebec plants, or if it was just a figment of this sample of data.

 

To determine if we should conclude that this pattern holds for the full population, conduct an appropriate hypothesis test for the hypotheses:

H0:μMiss−μQueb=0
Ha:μMiss−μQueb≠0
where μMiss represents the (unknown) true mean CO2 uptake of ALL Mississippi plants that are chilled at a concentration of 250 and μQueb represents the true (unkown) mean CO2 uptake of ALL Quebec plants that are chilled at a concentration of 250.

Report the test statistic, the parametric distribution being used for the test statistic (including degrees of freedom if appropriate), and p-value of the test.

```{r}
t.test(uptake ~ Type, 
       data = CO2.chilled.250,
       mu = 0,
       alternative = "two.sided",
       conf.level = .95)
```

Make a Q-Q Plot to determine if the distribution is normal 

```{r}
qqPlot(uptake ~ Type, data = CO2.chilled.250)
```

