---
title: "W4 Notes: Wilcoxon Tests"
output: 
  html_document: 
    theme: cerulean
    css: styles.css
---

```{r Libraries}
library(mosaic)
library(car)
```
```{r Example}

wilcox.test(length ~ sex,
            data=KidsFeet)
```


Since the wilcoxon test had ties in the rankings, let's try a t-test.

First test normality:
```{r}
qqPlot(length ~ sex, 
       data = KidsFeet)
```

Looks great! Now for the t-test:
```{r}
t.test(length ~ sex, 
       data = KidsFeet)

```
This works out great! Wilcoxon tests might not work well here because Wilcoxon is more used when you're analyizing ranks (like first, second, and third place in class rankings). 

If the t-test requirements are met, do that. If the distribution is not parametric, or if you're dealing with ordinal data instead of quantitative data. 


# Questions
What does "alternative hypothesis: true location shift is not equal to 0" mean?

- that you have repeat values in the list? This means that when ranking the values, there were ties between ranks.
- if there are a lot of ties, a Wilcoxon test won't work.
  

What is "stochastically" mean?

- when the distribution is taller, or wider than a normal distribution?
  


