---
title: "W6_class_notes"
output: html_document
---

```{r Prep}
library(tidyverse)
library(mosaic)
library(pander)
```

# Announcements
In 3 weeks (DO IT NOW)
- data set
- quesitn about the data set

best: find a small business that would like some help.
(KAI'S SHOE STORE? GAME PULSE? IMDB? SLEEP DATA? )

# Kruskal-Wallis Test

## Why
An ANOVA test requires the dataset with more than 2 factors to have a normal distribution. Kruskal-Wallis Tests can test more than 2 factors without needing a normal distribution.

Can also be used when data is ordered (ordinal) but does not have an exact measurement.

## What

## How
You really just do three things:
- run a kruskal
- make a boxplot
- show the datatable

### Example pigweights

```{r Dataset}
pigweights <- data.frame(Subject=1:56,
                   Litter=factor(rep(1:8, c(10,8,10,8,6,4,6,4))),
                   Weight=c(2.0,2.8,3.3,3.2,4.4,3.6,1.9,3.3,2.8,1.1,
                            3.5,2.8,3.2,3.5,2.3,2.4,2.0,1.6,
                            3.3,3.6,2.6,3.1,3.2,3.3,2.9,3.4,3.2,3.2,
                            3.2,3.3,3.2,2.9,3.3,2.5,2.6,2.8,
                            2.6,2.6,2.9,2.0,2.0,2.1,
                            3.1,2.9,3.1,2.5,
                            2.6,2.2,2.2,2.5,1.2,1.2,
                            2.5,2.4,3.0,1.4))
knitr::kable(pigweights, align='l')
```

```{r Kruskal}
kruskal.test(Weight ~ Litter, data=pigweights)
```
```{r Boxplot}
boxplot(Weight ~ Litter, data=pigweights, 
        col="lightgray", xlab="Litter Number", main="Weights of Piglets for 8 Different Litters from One Mother", ylab="Piglet Weight")
stripchart(Weight ~ Litter, data=pigweights, 
           vertical=TRUE, pch=16, add=TRUE, col="skyblue3")

```
```{r Datatable}
pander(favstats(Weight ~ Litter, data=pigweights)[,-10])
```

### Example Friendly

```{r Dataset}
library(car)
?Friendly
```
```{r Kruskal}
kruskal.test(
  correct ~ condition,
  data = Friendly
  )
```
We have insufficient evidence to reject the null.

```{r Boxplot}
boxplot(correct ~ condition,
  data = Friendly)
```
```{r Datatable}

```



