---
title: "W6_class_notes"
output: html_document
---

```{r Prep, include=FALSE}
library(tidyverse)
library(mosaic)
library(pander)
```

# Announcements
In 3 weeks (DO IT NOW)
- data set
- quesitn about the data set

best: find a small business that would like some help.
(KAI'S SHOE STORE? GAME PULSE? IMDB? SLEEP DATA? )

# Kruskal-Wallis Test

## Why
An ANOVA test requires the dataset with more than 2 factors to have a normal distribution. Kruskal-Wallis Tests can test more than 2 factors without needing a normal distribution.

Can also be used when data is ordered (ordinal) but does not have an exact measurement.

## What

## How
You really just do three things:
- run a kruskal
- make a boxplot
- show the datatable

### Example pigweights

```{r Pigweights Dataset}
pigweights <- data.frame(Subject=1:56,
                   Litter=factor(rep(1:8, c(10,8,10,8,6,4,6,4))),
                   Weight=c(2.0,2.8,3.3,3.2,4.4,3.6,1.9,3.3,2.8,1.1,
                            3.5,2.8,3.2,3.5,2.3,2.4,2.0,1.6,
                            3.3,3.6,2.6,3.1,3.2,3.3,2.9,3.4,3.2,3.2,
                            3.2,3.3,3.2,2.9,3.3,2.5,2.6,2.8,
                            2.6,2.6,2.9,2.0,2.0,2.1,
                            3.1,2.9,3.1,2.5,
                            2.6,2.2,2.2,2.5,1.2,1.2,
                            2.5,2.4,3.0,1.4))
knitr::kable(pigweights, align='l')
```

```{r Kruskal 1}
kruskal.test(Weight ~ Litter, data=pigweights)
```
```{r Boxplot 1}
boxplot(Weight ~ Litter, data=pigweights, 
        col="lightgray", xlab="Litter Number", main="Weights of Piglets for 8 Different Litters from One Mother", ylab="Piglet Weight")
stripchart(Weight ~ Litter, data=pigweights, 
           vertical=TRUE, pch=16, add=TRUE, col="skyblue3")

```
```{r Datatable 1}
pander(favstats(Weight ~ Litter, data=pigweights)[,-10])
```

### Example Friendly

```{r Friendly Dataset, include=FALSE}
library(car)
?Friendly
```
```{r Kruskal 2}
kruskal.test(
  correct ~ condition,
  data = Friendly
  )
```
We have insufficient evidence to reject the null.

```{r Boxplot 2}
boxplot(correct ~ condition,
  data = Friendly)
```
```{r Datatable 2}

```




# Skills Quiz
## Question 1

```{r set up, include=FALSE}
library(mosaic)

table(SaratogaHouses$fuel)
```

```{r, include=FALSE}
?SaratogaHouses

View(SaratogaHouses)
```

Say a homeowner in Saratoga County, New York is curious about whether upgrading their home from an oil heating fuel system to either a gas or electric fuel system would increase the resale value of their home. Use the SaratogaHouses dataset in R and a Kruskal-Wallis Test to answer the question, "which heating system results in the hightest distribution of home resale values (price)?"

$H_0$:  The null hypothesis for this study is that the prices of homes with either gas, oil, or electric heating fuel systems all come from the same distribution. 

$H_a$: The alternative hypothesis is that at least one type of fuel system results in a different distribution of prices of homes.

### Kruskal-Wallis Test 
Find test statistic, degrees of freedom, p-value, and highest median price of house for an individual fuel system.

```{r}
kruskal.test(
  price ~ fuel,
  data = SaratogaHouses
  )

boxplot(price ~ fuel, data=SaratogaHouses)

pander(
  favstats(
    price ~ fuel,
    data = SaratogaHouses
  )
  )
```


Make a conclusion

Reject the null. The highest median is for the gas fuel system.

## Question 2
Study the wages and pigweights examples for the Kruskal-Wallis Test before completing the following questions.

Copy and paste in the chi-squared results from those studies into the quiz.

## Question 3
Review the Kruskal-Wallis Test notes

## Question 4
Remember that customizing your table of contents is 5% of your grade. Work on it for this question and email your teacher a screen shot of what you did.

