---
title: "R Cheat Sheets & Notes"
output: 
  html_document: 
    theme: cerulean
    highlight: tango
    css: styles.css
---

## Cheat Sheets

* [R Colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

* [R Base Graphics Cheat Sheet](http://www.joyce-robbins.com/wp-content/uploads/2016/04/BaseGraphicsCheatsheet.pdf)

* [ggplot2 Intermediate Graphics Cheat Sheet](https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

* [R Base Commands Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2016/10/r-cheat-sheet-3.pdf)

* [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

* [Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts)



* [More Cheat Sheets by Rstudio](https://rstudio.com/resources/cheatsheets/)

## Notes

Add your own notes here...

```{r}
# Add R commands you want to remember here...

```

```{r}
# Or create other R chunks to put them in like this...

```

```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.
```

Datacamp.com made me aware that ggplot can scale the x and/or y axis on a logorithmic scale. Since I don't see it elsewhere in the notebook I'll add it here for now.

```{r installs, eval=FALSE}
install.packages("gapminder")
```


```{r libraries}
library(gapminder)
library(ggplot2)
library(dplyr)
```


```{r scale_x_log10}
#filter
gapminder_2007 <- gapminder %>% 
  filter(year == 2007)

#plot
ggplot(gapminder_2007,
       aes(x = gdpPercap,
           y = lifeExp,
           color = continent)
       ) + geom_point() + scale_x_log10()
```

