---
title: "My Simple Logistic Regression"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---
<hr />

<!-- For the second draft of this analysis, consider filtering out all of the $0 revenue values as well. -->





<!-- for this analysis if you follow the outline of the challenger example you'll get an automatic 100% -->

<!-- Find some data of interest to you that can be analyzed with a logistic regression. Complete your Logistic Regression Analysis using this file. 

Hint, run the command 

> data() 

in your Console to see available datasets in R that you could use for this Analysis. You are also welcome to collect data yourself, or find other available datasets online for this Analysis. Just be careful to not use anyone else's "analysis" for this Analysis. Also, no conducting surveys unless you have IRB approval to do so. -->

```{r, include=FALSE}
library(mosaic)
library(ResourceSelection) 
library(DT)
setwd("C:/Users/BAG20/Desktop/Data Analytic & School/R-Python-Statistical-Analysis-Notebook/Analyses/Logistic Regression")
```

```{r}
tmdb <- read.csv("../../Data/tmdb_5000_movies.csv", header=TRUE)
# head(movies)
# str(movies)
# summary(movies)
movies <- ggplot2movies::movies
```


#### Background

Making movies is a significant financial venture. When millions, and even billions, of dollars are at stake, it is very important to know if people will like the movie. We will begin an analysis on this issue by finding out what budget amounts are more likely to result in a movie that generates more revenue than was spent in the budget. 



```{r}
#load csv
tmdb <- read.csv("../../Data/tmdb_5000_movies.csv", header=TRUE)

#select relevant columns
tmdb2 <- tmdb %>% 
  select(
    original_title,
    budget,
    revenue
    ) %>% 
  mutate(
    profitable = case_when(
      revenue > budget ~ 1,
      revenue <= budget ~ 0
    )
  )

#identify movies of the fantasy genre
#tmdb2$fantasy <- grepl("Fantasy", tmdb2$genres, fixed = TRUE)
```
Due to this data being compiled via web scrapping, I will assume that the movies that have a record budget of 0 in combination with a revenue of 0 should be treated as if the 0 values where actual NA's. Those movies will be filtered out of this analysis as a result.

```{r}
#filter out NA's and erroneous data
tmdb3 <- tmdb2 %>% 
  filter(tmdb2$budget & tmdb2$revenue == 0)

datatable(tmdb3)  
```

#### Logistic Model

The model for the logistic regression model for the probability of a movie being profitable based on budget is as follows:

$$
  P(Y_i = 1|x_i) = \frac{e^{\beta_0+\beta_1 x_i}}{1+e^{\beta_0 + \beta_1 x_i}} = \pi_i
$$

where for observation $i$: 

* $Y_i = 1$ denotes that the budget was less than the revenue earned, 
* $Y_i=0$ denotes that the budget was equal to or less than the revenue gained, and 
* $x_i$ denotes the movies budget. 

Note that if $\beta_1$ is equal to zero in the above model, then it doesn't matter what $x_i$ (budget) is equal to since $\beta_1 x_i$ will always equal 0. So when it comes to probability, we could test the hypothesis that

$$
  H_0: \beta_1 = 0 \\
  H_a: \beta_1 \neq 0
$$

<br />

#### Fitting the Model

Now, we'll find what $\beta_0$ and $\beta_1$ are estimated to be for the `tmdb` data.

```{r}
tmdb.glm <- glm(profitable > 0 ~ budget, 
                data = tmdb3, 
                family = binomial)
summary(tmdb.glm)
```

Thus the estimated model for $\pi_i$ is given by
$$
  P(Y_i = 1|x_i) \approx \frac{e^{-2.657e^{1}- 7.914e^{22} x_i}}{1+e^{-2.657e^{1} - 7.914e^{22}}} = \hat{\pi}_i
$$
where $b_0 = -2.657e^{1}$ is the value of the `(Intercept)` and $b_1 = 7.914e^{22}$ is the value of `budget`. `(Intercept)` estimates $\beta_0$ and `budget` estimates $\beta_1$.

#### Conclusion

Since the $p$-value for the test of `budget` shows an insignificant result $(p = 1.000)$ there is insufficient evidence to conclude that $\beta_1 \neq 0$. We accept the null hypothesis that budget does not affect the probability of a movie making more money than was budgeted.

<br />

#### Visualizing the Model

Simple logistic regressions can be visualized like follows:

```{r}
plot( profitable > 0 ~ budget,
      data = tmdb3, 
      main = "",
      ylab = 'Probability of a Movie being Profitable',
      pch = 16)

curve(exp((-2.657*exp(1))-(7.914*exp(22))*x)/(1+exp((-2.657*exp(1))-(7.914*exp(22))*x)),
      from = min(tmdb3$budget),
      to = max(tmdb3$budget),
      add = TRUE)
```


#### Diagnosing the Model

Lets test if the logistic regression was a good fit for the data. Since there are many replicated $x$-values, we'll use the deviance goodness of fit test.

```{r Deviance Goodness of Fit Test}
pchisq(
  3.1154e-09, #residual deviance
  535, #df for residual deviance
  lower.tail = FALSE
)
```

Recall that the null hypothesis is that the logistic regression is a good fit for the data, thus we reject the null, and claim that the logistic regression is not appropriate.

#### Interpreting the Model

The value of $e^{b_0} = -2.657e^{1} = -26.57$ which shows that the odds for the movie to be provitable when the budget is $\$0$ or ($Y_i=1 \text{ when } x_i = 0$) is -26.57. This is uninterpretable since that would equal -2657% chance and we can't go below -100%. 

The value of $e^{b_0} = 7.914e^{-22}$ which shows that the odds of the movie being profitable increases by a factor of about 7.914 divided by about a fifth of a googol; a number virtually equivalent to 0.