---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(lmtest)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
late <- read.csv("../../Data/late.csv", header=TRUE)
# Then type:
# > View(late)
# in your R Console
```



----

<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

**Comments to Critiquers:** 


<!--End your comments to your critiquers before this line.-->

----


## Background {.tabset .tabset-fade}

### Overview
<!-- Be sure to add your own background here before beginning the Analysis. -->
This report will attempt to measure the effect tardiness fines have on daycare pick up times.

The data comes from ucsd.edu and the details of their study is quoted in the "Study Details" tab.






<!--End your background before here.-->
<!--Click the little black arrow to 
the left to expand/hide these next sections.-->
### Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather than to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>


### The Data (Wide)


The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


### The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
# This code reshapes the data into "long" format called Late.
# To get the "Late" dataset into your R Console, 
# you need to click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 
# Then type:
#   > View(Late)
# in your R Console
Late <- reshape(late,
                varying = paste("Week",1:20, sep=""), 
                v.names = "No.ofLateChildren",
                timevar = "Week", 
                times = 1:20, 
                idvar = "Center",
                new.row.names = 1:200,
                direction = "long")
pander(Late)
```
```{r}
Late$Period <- cut(Late$Week, c(0, 4, 16, 20), labels = c("Pre-Fine", "Fine", "Postfine")) 
```


 
##  
<!-- You can begin writing again after here.-->



## Analysis

### Hypothesis

There are two recorded factors that could possibly affect how many parents are late: Treatment and Period. The Treatment factor has two levels, *Control* and *Fine*. The Period factor has three levels, *Pre-Fine*, *Fine*, and *PostFine*.

For a two-way ANOVA there are three sets of hypotheses, one set for each question we can ask:

#### Question 1: Does the Treatment factor affect the number of children picked up late?
$$
  H_0: \mu_{Control} = \mu_{Fine} = \ldots = 0
$$
$$
  H_a: \mu_i \neq 0 \ \text{for at least one} \ i
$$

#### Question 2: Does the Period factor affect the number of children picked up late?
$$
  H_0: \mu_{Pre-Fine} = \mu_{Fine} = \mu_{Postfine} = \ldots = 0
$$
$$
  H_a: \mu_j \neq 0 \ \text{for at least one} \ j
$$

#### Question 3: Does the effect of the Treatment factor change for different periods?
$$
  H_0: \text{The effect of} \ Treatment \ \text{is the same for all levels of} \ Period.
$$
$$
  H_a: \text{The effect of} \ Treatment \ \text{is different for at least one levels of} \ Period.
$$


The mathematical model is as follows and will be used to conduct an ANOVA:

$$
  Y_{ijk} = \mu + \alpha_i + \beta_j + \alpha\beta_{ij} +\epsilon_{ijk} \ 
$$
where $\mu$ is the grand mean, $\alpha$ is the Treatment factor with levels Fine = 1 and Control = 2, $\beta_j$ is the Period factor with levels Pre-Fine = 1, Fine = 2, and Postfine = 3,  $\alpha\beta_{ij}$ is the interaction of the two factors which has 2 x 3 = 6 levels, and $\epsilon_{ijk}\sim N(0,\sigma^2)$ is the error term.


### ANOVA

I will conduct an ANOVA with the following code:

```{r ANOVA}
Late.aov <- aov(
  No.ofLateChildren ~ Treatment + Period + Treatment:Period,
  data = Late
  )

summary(Late.aov) %>% 
  pander()
```

The ANOVA returns low p-values for each of the three hypotheses. Therefore, at least one level for each of the factors has a different average amount of tardiness then from the other levels of that factor.

In other words, each of the factors has an effect on how many children are picked up late.

I'll show the results of the ANOVA graphically by plotting each factor.


```{r Plot Demonstrating Treatment}
xyplot(
  No.ofLateChildren ~ Treatment, 
  data = Late, 
  main = "Plot Demonstrating Treatment", 
  jitter.x=TRUE,
  pch=16)
```
```{r Plot Demonstrating Period}
xyplot(
  No.ofLateChildren ~ Period, 
  data = Late, 
  main = "Plot Demonstrating Period", 
  jitter.x=TRUE,
  pch=16)

```
```{r Plot Significance of the Interaction}
xyplot(
  No.ofLateChildren ~ Period,
  data = Late, 
  groups = Treatment,
  type=c("p","a"),
  main="Significance of the Interaction",
  auto.key=list(corner=c(1,1)))
```

Here we can see that the fined group had more children be picked up late over the course of the experiment while the control group had either insignificant change or a minor decrease in the amount of children picked up late.

### Requirements Check

I'll now check the requirements for the two-way ANOVA.

```{r}
par(mfrow = c(1,2))
plot(Late.aov, 
     which = 1:2,
     col = "skyblue")
```

The sample of data can be considered to be representative of the population since there are more than 30 observations.

As shown, the data is normally distributed in each group and the population variance of each group can be assumed to be the same since the *residuals* from the ANOVA show a constant variance.


## Conclusion
For whatever reason, it looks like both the presence of a fine and the removal of the fine increased the number of children that were picked up late from the private day care. There could be many reasons for this, but we'd need more data to be sure. In the meantime, private daycares in India should refrain from fining parents that arrive late to pick up their children, as not doing so seems to increase the number of children picked up late each week from about 10 to about 20 children in maximum capacity day cares.
