---
title: "Recalling Words"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r load libraries, include=FALSE}
library(mosaic)
library(car)
library(pander)
library(DT) #You may need to run: install.packages("DT") 
```

```{r examine data, eval=FALSE}
# Play the chunk above and this one to get the data into your Console
View(Friendly)
?Friendly
```


<br />

#### Background

Many teachers and other educators are interested in understanding how to best deliver new content to students. In general, they have two choices of how to do this.

1. The Meshed Approach
    * Deliver new content while simultaneously reviewing previously understood content.

2. The Before Approach
    * Deliver new content after fully reviewing previously understood content.

A study was performed to determine whether the *Meshed* or *Before* approaches to delivering content had any positive benefits on memory recall. 

<div style="padding-left:15px;">

## <a href="javascript:showhide('uniquename')">The Experiment <span style="font-size:8pt;">(click to view)</span></a>


<div id="uniquename" style="display:none;">

Individuals were seated at a computer and shown a list of words. Words appeared on the screen one at a time, for two seconds each, until all words had been shown (40 total). After all words were shown, they were required to perform a few two-digit mathematical additions (like 15 + 25) for 15 seconds to avoid immediate memory recall of the words. They were then asked to write down as many of the 40 words as they could remember. They were given a maximum of 5.3 minutes to recall words.

The process of showing words and recalling words was repeated four times with the same list of words each time (four chances to get it right). The presentation of the first trial was the same for all treatment conditions. However, trials 2, 3, and 4 were slightly different for each treatment condition.

<div style="padding-left:15px;">

The `SFR` group (the control group) stands for Standard Free Recall. In all four trials the same list of 40 words was presented, in a random order each time.

The `Before` group also used the same 40 words during each trial. However, any words that were correctly recalled in a previous trial were presented first, or *before* the words that were not recalled in the last trial. After all the correct words were presented in random order, the non-recalled words were presented in a random order.

The `Meshed` group also used the same 40 words during each trial. However, words that were correctly recalled in a previous trial were alternated with a missed word during the next presentation order. 

</div>

The data records the number of correctly recalled words (out of the 40 possible) from the fourth trial. Results were obtained for 30 students, 10 in each of the three treatment groups: `SFR`, `Before`, and `Meshed`. 

</div>

##### <a href="javascript:showhide('uniquename2')">The Data <span style="font-size:8pt;">(click to view)</span></a>

<div id="uniquename2" style="display:none;">

The results from the study can be found in the `Friendly` data set in R after loading `library(car)`. 

Click the "Code" button to see the data.


```{r view raw data}
datatable(Friendly, options=list(lengthMenu = c(3,10,30)))

```


</div>
</div>

<br />

## Analysis

### Research Question
<!-- Note that your goal is to use the Friendly data to show whether or not the Meshed or Before methods have any positive benefit on memory recall. -->

We'll be testing to see if either of the two treatment groups, the ones that used the "Meshed" or "Before" methods, had a greater positive affect on memory recall than control group, the one that used the "SFR" method. 

This will require two indendent samples tests: Meshed compared to SFR and Before compared to SFR. Before we can determine what type of test we can do, we'll need to check the normality of our data.

### Describe the Data

We'll create a six number summary and a Q-Q plot of each of the three samples.

```{r Check for Normality}
Friendly %>% 
  group_by(condition) %>% 
  summarise(min = min(correct),
            Q1 = quantile(correct, 0.25),
            mean = mean(correct), 
            median = median(correct), 
            Q3 = quantile(correct, 0.75),
            max = max(correct),
            "Sample Size" = n()
            ) %>%
  pander()

qqPlot(correct ~ condition,
       data = Friendly)

stripchart(correct ~ condition,
           data = Friendly, 
           method = "stack")
```

From these we can see that the data is not normal, none of the samples are large enough to be considered normal, and the distributions are not identically distributed. We'll have to conduct a Wilcoxon Rank Sum Test.

### Make inferences

Now that we've determined what type of test we'll conduct, I'll set the official hypothesis.

The null hypothesis is that the difference in the treatment and control samples' medians are equal to zero.

The alternative hypothesis is that the difference in the treatment and control samples' medians are not equal to zero.

Our alpha is equal to 0.05.

The Wilcoxon Rank Sum Tests are as follows:

```{r Wilcoxon Rank Sum Test 1}
wilcox.test(
  Friendly$correct[Friendly$condition == "SFR"],
  Friendly$correct[Friendly$condition == "Before"],
  mu = 0,
  alternative = "greater",
  conf.level = 0.95
  
)
```

```{r Wilcoxon Rank Sum Test 2}
wilcox.test(
  Friendly$correct[Friendly$condition == "SFR"],
  Friendly$correct[Friendly$condition == "Meshed"],
  mu = 0,
  alternative = "greater",
  conf.level = 0.95
)
```

Since the p-value for the "Before" method hypotheses is _, we have _sufficient evidence to reject the null hypothesis for the "Before" method. 

Since the p-value for the "Meshed" method hypotheses is _, we have _sufficient evidence to reject the null hypothesis for the "Meshed" method.

### Conclusion
