---
title: "Chi Squared Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


<!-- Directions

Come up with a chi squared test analysis of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You can use data from 
    - An R data set
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - Do not violate the "Human Subjects Rules" of the university: http://www.byui.edu/executive-strategy-and-planning/institutional-research/research-and-assessments/research-and-surveys

* You need a sample size large enough to meet the requirements of the chi squared test. Unfortunately, the sample size is completely dependent on the situation you select. A good rule of thumb is 10 observations per "combination-group" in your data.

-->


## Background
Social issues are often hard to quantify but detailed surveys yield great insites if the sampling method is done currectly.

The university of Berkeley has a large dataset of surveys available for viewing at sda.berkelyey.edu ( [clickhere](http://sda.berkeley.edu/sdaweb/analysis/?dataset=gss12) ).

Their 2012 ISSP Module contains a survey question about worklife balance and I am curious to see if there is any correlations between a respondents' gender and how they answered the question. 


Respondents' gender was classified as either male or female. 

The respondents were asked to respond to the statement, "It has been difficult for you to fulfil your family responsibilities because of the amount of time you spent on your job." Their answers could be any of the following:

0: Doesn't apply/no job
1: Several times a Week
2: Several Times a Month
3: Once or Twice
4: Never
8: Don't Know
9: No answer


## Hypothesis
Our official questions is: Does time spent working have varing impact on each genders ability to fulfilling family responsibilities. Our null and alternative hypothesis are stated below:

$H_0 : \text{The respondents' sex and their difficult to fulfil their family responsibilities are independent}$
$H_a : \text{The respondents' sex and their difficult to fulfil their family responsibilities are associated (i.e. not independent)}$

We will use $\alpha = 0.05$ as our significance level.

## Analysis
First lets make a table of the counts for each answer by gender. I would assume that due to culture most of the female respondents will have answered 0: N/A.
<!-- SEX & JOBVSFA1 -->
```{r prepare environment}
library(readr)
GSS2012 <- read_delim("../../Data/GSS2012.csv", 
                      "\t", escape_double = FALSE, trim_ws = TRUE)

library(mosaic)
rawdata <- GSS2012 %>%  
  select(sex, JOBVSFA1)
  

# Filter out the IAP results
filt.data <- rawdata %>% 
  filter(JOBVSFA1 != -1)

# Reformate for visability
data <- filt.data %>% 
  mutate(JobvsFam = JOBVSFA1) %>% 
  select(sex, JobvsFam) %>% 
  mutate(
    sex = case_when(
      sex == 1 ~ "Male",
      sex == 2 ~ "Female"
    ),
    JobvsFam = case_when(
      JobvsFam == 0 ~ "0: N/A",
      JobvsFam == 1 ~ "1: Weekly",
      JobvsFam == 2 ~ "2: Monthly",
      JobvsFam == 3 ~ "3: Once or Twice",
      JobvsFam == 4 ~ "4: Never",
      JobvsFam == 8 ~ "8: Don't Know",
      JobvsFam == 9 ~ "9: No answer"
    )
  )
```

```{r table of the data}
jobtable <- table(data$JobvsFam, data$sex)
jobtable2 <- t(jobtable)

jobtable
```

It does look like a largest group is female respondents stating that the question is not applicable. Let's visualize this with a barplot.

```{r barplot of the data}
barplot(jobtable2, 
        beside = TRUE, 
        col = c("pink","blue"),
        horiz = TRUE,
        las = 2,
        legend.text = TRUE,
        args.legend =
          list(x = "topright",
               bty = "n",
               title = "Gender"),
        xlab = "Response",
        main = "Respondents Work-Life balance")
```

<!-- intro text -->
```{r chi-squared test of the data}

```

<!-- intro text -->
```{r expected values table}

```
<!-- are the requirements met? -->
<!-- do we accept or reject the hypothesis? -->

<!-- intro text -->
```{r table of residuals}

```
<!-- explain residuals -->

## Interpretation
<!-- text -->

