---
title: "Chi Squared Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


<!-- Directions

Come up with a chi squared test analysis of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You can use data from 
    - An R data set
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - Do not violate the "Human Subjects Rules" of the university: http://www.byui.edu/executive-strategy-and-planning/institutional-research/research-and-assessments/research-and-surveys

* You need a sample size large enough to meet the requirements of the chi squared test. Unfortunately, the sample size is completely dependent on the situation you select. A good rule of thumb is 10 observations per "combination-group" in your data.

-->


## Background
Social issues are often hard to quantify but detailed surveys yield great insights if the sampling method is done correctly.

The university of Berkeley has a large dataset of surveys available for viewing at sda.berkelyey.edu ( [clickhere](http://sda.berkeley.edu/sdaweb/analysis/?dataset=gss12) ).

Their 2012 ISSP Module contains a survey question about work life balance and I am curious to see if there are any correlations between a respondents' gender and how they answered the question. 


Respondents' gender was classified as either male or female. 

The respondents were asked to respond to the statement, "It has been difficult for you to fulfil your family responsibilities because of the amount of time you spent on your job." Their answers could be any of the following:

0: Doesn't apply/no job
1: Several times a Week
2: Several Times a Month
3: Once or Twice
4: Never
8: Don't Know
9: No answer


## Hypothesis
Our official question is: Does time spent working have varying impact on each genders ability to fulfilling family responsibilities. Our null and alternative hypothesis are stated below:

$H_0 : \text{The respondents' sex and their difficult to fulfil their family responsibilities are independent}$
$H_a : \text{The respondents' sex and their difficult to fulfil their family responsibilities are associated (i.e. not independent)}$

We will use $\alpha = 0.05$ as our significance level.

## Analysis

First let’s make a table of the counts for each answer by gender. I would assume that due to culture most of the female respondents will have answered 0: N/A.
<!-- SEX & JOBVSFA1 -->
```{r prepare environment}
library(readr)
GSS2012 <- read_delim("../../Data/GSS2012.csv", 
                      "\t", escape_double = FALSE, trim_ws = TRUE)

library(mosaic)
rawdata <- GSS2012 %>%  
  select(sex, JOBVSFA1)
  

# Filter out the IAP results
filt.data <- rawdata %>% 
  filter(JOBVSFA1 != -1)

# Reformate for visibility
data <- filt.data %>% 
  mutate(JobvsFam = JOBVSFA1) %>% 
  select(sex, JobvsFam) %>% 
  mutate(
    sex = case_when(
      sex == 1 ~ "Male",
      sex == 2 ~ "Female"
    ),
    JobvsFam = case_when(
      JobvsFam == 0 ~ "0: N/A",
      JobvsFam == 1 ~ "1: Weekly",
      JobvsFam == 2 ~ "2: Monthly",
      JobvsFam == 3 ~ "3: Once or Twice",
      JobvsFam == 4 ~ "4: Never",
      JobvsFam == 8 ~ "8: Don't Know",
      JobvsFam == 9 ~ "9: No answer"
    )
  )
```

```{r table of the data}
jobtable <- table(data$JobvsFam, data$sex)
jobtable2 <- t(jobtable)

jobtable
```

It does look like a largest group is female respondents stating that the question is not applicable. Let's visualize this with a bar plot.

```{r bar plot of the data}
barplot(jobtable2, 
        beside = TRUE, 
        col = c("pink","blue"),
        horiz = TRUE,
        las = 2,
        legend.text = TRUE,
        args.legend =
          list(x = "topright",
               bty = "n",
               title = "Gender"),
        xlab = "Response",
        main = "Respondents Work-Life balance")
```

Now that we've examined and prepared the data, I will conduct a Chi-Squared Test. This is how we'll know if we should reject or fail to reject the null hypothesis.
```{r Pearson’s Chi-squared test}
chi <- chisq.test(jobtable)

library(pander)
pander(chi)

pander(chi$expected)
```

These results tell us that though there are expected values that or less than 5, the average is greater than 5 and the requirements for the test have been met.

Our $p$-value (9.005e-06) is less than $\alpha$ so we reject the null hypothesis and conclude that the work-life balance and gender of respondents are associated.

In order to get a clearer picture of how these two factors are related, I have included a chart of how far our observed values where from the expected values.
```{r table of residuals}
pander(chi$residuals)
```

The only major differences came with the observed counts of respondents answering "0:N/A" or "4: Never". 

## Interpretation

Something interesting to note is that for every question other than "0: N/A" or "9: No answer" the female respondents had less counts than expected. Male respondents likewise had more counts than expected in every area other than those same two answers. In the case of "9: No answer", female respondents were barely above the expected values and male respondents were barely below the expected count. 

It seems that that an unusually high amount of women are unemployed and thus don't have work-life balancing issues, but at the same time there are an unusually low amount of women who answered that they never had work-life balance issues. Likewise, and unusually low amount of men were unemployed, and an unusually high amount of men said that work never interfered with their duties at home.

This might indicate a reverse correlation between employment and work-life balance: it may be that as a group's employment rate increases the amount of times their work interrupts home duties decreases. In the case of women, their work-life balance is worse than expected and for men the balance is better than expected. 

<!-- This might indicate a type of polarization on this issue: either Women were unemployed, or their work responsibilities interrupted their responsibilities at home to some degree. The reverse being the trend for men: either men had never had their work responsibilities  -->

